---
title: "Challenge 02: Revenue Feature Engineering & Modeltime Workflow"
subtitle: "DS4B 203-R, Time Series Forecasting for Business"
author: "Business Science"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = F,
    warning = F,
    paged.print = FALSE, 
    # This should allow Rmarkdown to locate the data
    root.dir = rprojroot::find_rstudio_root_file()
)
```

# Challenge Objective

Your goal is to perform an 8-week revenue forecast. You'll need to engineer features that help. In this challenge, you'll:

1. Add event data to the transactions revenue dataset 
2. Preprocess the data
3. Create multiple `recipes`
3. Implement a `modeltime` forecasting workflow using multiple linear regression workflow objects


# Libraries

```{r}
# Modeling
library(tidymodels)
library(modeltime)

# Core
library(tidyverse)
library(timetk)
library(lubridate)
library(janitor)
```


# Collect Data

Read in the following data sets. 

## Transactions Revenue

```{r}
transactions_tbl  <- read_rds("../00_data/transactions_weekly.rds")
transactions_tbl
```

## Product Events

The product events data consists of date and events that are known to affect sales. 

```{r}
product_events_tbl <- read_rds("../00_data/product_events.rds")
product_events_tbl
```

# Preparing Data

Our objectives are to:

- Aggregate data to common time-stamps
- Apply any transformations
- Detect any lags & add rolling features
- Create a Full Data Set: Adding Future observations, lags, and external regressors


## Aggregate Revenue by Week

1. Start with `transactions_weekly_tbl`
2. Use `summarise_by_time()` with `.by = "week"`, and `sum()` the revenue.
3. Save as a new variable called `transactions_weekly_tbl`

```{r}
# YOUR CODE HERE
transactions_weekly_tbl <- transactions_tbl %>%
    summarise_by_time(
        .date_var = purchased_at, .by = "week", revenue = sum(revenue)
    )
```



## Aggregate Events by Week

1. Start with `product_events_tbl`
2. Use `add_column()` to create a column called "event_val". Set the values in the column to `1`. 
3. Use `group_by()` to group on the "event" column.
4. Use `summarise_by_time()` with `.by = "week"`, and `sum()` the "event_val" column.
5. Ungroup the data. 
6. Pivot the data wider so we have a "date" column:
    - Use: `names_from   = event`
    - Use: `values_from  = event_val`
    - Use: `values_fill  = 0`
    - Use: `names_prefix = "event_"`
7. Clean the names with `janitor::clean_names()`
7. Save as a new variable called `product_events_weekly_tbl`

```{r}
# YOUR CODE HERE
product_events_weekly_tbl <- product_events_tbl %>%
    add_column(event_val = 1) %>%
    group_by(event) %>%
    summarise_by_time(
        .date_var   = date
        , .by       = "week"
        , event_val = sum(event_val)
    ) %>%
    ungroup() %>%
    pivot_wider(
        names_from = event
        , values_from = event_val
        , values_fill = 0
        , names_prefix = "event_"
    ) %>%
    clean_names()
```

# Visualizations

## Visualize Revenue

Use `plot_time_series()` to visualize the revenue. 

- Look for outliers & any data issues
- Try out a `log()` transformation to see the effect on the time series

```{r}
# YOUR CODE HERE
transactions_weekly_tbl %>%
    plot_time_series(
        .date_var = purchased_at
        , .value  = log(revenue)
    )
```

## Visualize ACF

Visualize the ACF using `plot_acf_diagnostics()` using a `log()` transformation. Look for:

- Any frequencies we can include?
- Any lags we can include? (Hint - What is our forecast horizon?)

```{r}
# YOUR CODE HERE
transactions_weekly_tbl %>%
    plot_acf_diagnostics(
        .date_var = purchased_at
        , .value  = log(revenue)
    )
```


## Log-Standardize Revenue (Target)

- Start with `transactions_weekly_tbl`
- Apply log-standardization:
    - Apply Log transformation using `log()`
    - Apply standardization to mean = 0, sd = 1 using `standardize_vec()`
- Store the resulting data as `transactions_trans_weekly_tbl`

```{r, message = TRUE}
# YOUR CODE HERE
transactions_trans_weekly_tbl <- transactions_weekly_tbl %>%
    mutate(revenue = log(revenue)) %>%
    mutate(revenue = standardize_vec(revenue))
```

Save the mean and standard deviation as `std_mean` and `std_sd`. We'll need this for inverting.

```{r}
# YOUR CODE HERE
std_mean = 11.2910425517621
std_sd   = 0.655408027721517
```


Visualize the log-standardized transactions using `plot_time_series()`. This confirms the transformation was performed successfully. 

```{r}
# YOUR CODE HERE
transactions_trans_weekly_tbl %>%
    plot_time_series(
        .date_var = purchased_at
        , .value  = revenue
    )
```

# Create Full Data Set

```{r}
# transactions_trans_weekly_tbl %>% write_rds("challenge_02_data_checkpoints/transactions_trans_weekly_tbl.rds")
# product_events_weekly_tbl %>% write_rds("challenge_02_data_checkpoints/product_events_weekly_tbl.rds")
# list(std_mean = std_mean, std_sd = std_sd) %>% write_rds("challenge_02_data_checkpoints/standardize_params.rds")

# Checkpoint data
product_events_weekly_tbl     <- read_rds("challenge_02_data_checkpoints/product_events_weekly_tbl.rds")
transactions_trans_weekly_tbl <- read_rds("challenge_02_data_checkpoints/transactions_trans_weekly_tbl.rds")
standardize_params            <- read_rds("challenge_02_data_checkpoints/standardize_params.rds")

std_mean <- standardize_params$std_mean
std_sd   <- standardize_params$std_sd
```


We'll use these parameters to create our "full dataset". We've selected an 8-week forecast horizon. Our lag period is 8 weeks and we'll try out a few rolling averages at various aggregations. 

```{r}
horizon         <- 8
lag_period      <- 8
rolling_periods <- c(4, 8, 16, 24, 36)
```


## Prepare the full data

Next, join the aggregated weekly transactions revenue data and the product events data. 

1. Start with `transactions_weekly_tbl`
2. __Add the future window:__ Use `bind_rows()` and `future_frame()` to extend the data frame `.length_out = horizon`.
3. __Add autocorrelated lags:__ Use `tk_augment_lags()` to add a `.lags = lag_period`
4. __Add rolling features from our lag__: Use `tk_agument_slidify()` to add `.period = rolling_periods`. Use `mean` as the rolling function. Make sure to "center" with "partial" windows. 
5. __Add events__:
    - Left join `product_events_weekly_tbl`
    - Fill in the missing values with zero for any column that start with "event_"
6. Rename any columns that contain "lag". Modify to start with "lag_"
6. Save the output as `full_tbl`.


```{r}
# YOUR CODE HERE
 full_tbl <- transactions_weekly_tbl %>%
    bind_rows(
        future_frame(., .date_var = purchased_at, .length_out = horizon)
    ) %>%
    tk_augment_lags(.value = revenue, .lags = lag_period) %>%
    tk_augment_slidify(
        .value = revenue
        , .period = rolling_periods
        , .f = mean
        , .align = "center"
        , .partial = TRUE
    ) %>%
    left_join(
        product_events_weekly_tbl
        , by = c("purchased_at"="date")
    ) %>%
    mutate(across(starts_with("event_"), .fns = ~ ifelse(is.na(.), 0, .))) %>%
    rename_with(.cols = contains("lag"), .fn = ~ str_c("lag_", .))
```

## Visualize the Full Data

Visualize the features, and review what you see. 

1. Start with `full_tbl`
2. `pivot_longer` every column except "purchased_at"
3. Use `plot_time_series()` to visualize the time series coloring by "name". 

Review the visualization selecting one feature at a time and answering the following questions:
    
    - Do the rolling lags present any issues? 
    - Which rolling lag captures the trend the best?
    - Do you expect either of the Product Events features to help?

```{r}
# YOUR CODE HERE
full_tbl %>%
    pivot_longer(
        -purchased_at
    ) %>%
    plot_time_series(
        .date_var = purchased_at
        , .value = value
        , .smooth = FALSE
        , .color_var = name
    )
```

# Model Data / Forecast Data Split

```{r}
# full_tbl %>% write_rds("challenge_02_data_checkpoints/full_tbl.rds")

# Checkpoint data
full_tbl <- read_rds("challenge_02_data_checkpoints/full_tbl.rds")
```

Create a `data_prepared_tbl` by filtering `full_tbl` where "revenue" is non-missing. 

```{r}
# YOUR CODE HERE
data_prepared_tbl <- filter(full_tbl, !is.na(revenue))
```

Create a `forecast_tbl` by filtering `full_tbl` where "revenue" is missing. 

```{r}
# YOUR CODE HERE
forecast_tbl <- filter(full_tbl, is.na(revenue))
```


# Train / Test Split

```{r}
# data_prepared_tbl %>% write_rds("challenge_02_data_checkpoints/data_prepared_tbl.rds")
# forecast_tbl %>% write_rds("challenge_02_data_checkpoints/forecast_tbl.rds")

# Checkpoint data
data_prepared_tbl <- read_rds("challenge_02_data_checkpoints/data_prepared_tbl.rds")
forecast_tbl      <- read_rds("challenge_02_data_checkpoints/forecast_tbl.rds")
```


## Split into Train / Test Sets

- Start with `data_prepared_tbl`
- Use `time_series_split()` to create a single time series split. 
    - Set `assess = horizon` to get the last 8-weeks of data as testing data. 
    - Set `cumulative = TRUE` to use all of the previous data as training data. 
- Save the object as `splits`

```{r}
# YOUR CODE HERE
splits <- data_prepared_tbl %>%
    time_series_split(
        .date_var    = purchased_at
        , assess     = horizon
        , cumulative = TRUE
    )
```



# Feature Engineering

```{r}
# write_rds(splits, "challenge_02_data_checkpoints/splits.rds")

# Checkpoint data
splits <- read_rds("challenge_02_data_checkpoints/splits.rds")
```

## Create a Preprocessing recipe

Make a preprocessing recipe using `recipe()`. Note - It may help to `prep()` and `juice()` your recipe to see the effect of your transformations. 

- Start with `recipe()` using "revenue ~ ." and `data = training(splits)`
- Add the following steps:
    - `step_timeseries_signature()` using the date feature
    - Remove any newly created features that:
        - Contain ".iso"
        - End with "xts"
        - Contain "day", "hour", "minute", "second" or "am.pm" (because this is a weekly dataset and these features won't add any predictive value)
    - Normalize all numeric data except for "revenue" (the target) with `step_normalize()`.
    - Dummy all categorical features with `step_dummy()`. Set `one_hot = TRUE`.
    - Add a fourier series at periods 4 and 20. Set K = 2 for both. 

```{r}
# YOUR CODE HERE
recipe_spec_base <- recipe(revenue ~ ., data = training(splits)) %>%
    step_timeseries_signature(purchased_at) %>%
    step_rm(matches("(.iso)|(xts)|(_day)|(hour)|(minute)|(second)|(am.pm)")) %>%
    step_normalize(
        !matches("(revenue)|(purchased_at)|(_revenue)")
        , all_numeric()
    ) %>%
    step_dummy(all_nominal(), one_hot = TRUE) %>%
    step_fourier(purchased_at, period = c(4, 20), K = 2)
```


# Modeling

```{r}
# write_rds(recipe_spec_base, "challenge_02_data_checkpoints/recipe_spec_base.rds")

# Checkpoint data
recipe_spec_base <- read_rds("challenge_02_data_checkpoints/recipe_spec_base.rds")
```


## Spline Model

### Visualize

Use `plot_time_series_regression` to test out several natural splies:

- Use .formula to try out `splines::ns()` with degrees of freedom 1, 2, 3, and 4. 

Which value of `df` would you select?

```{r}
# YOUR CODE HERE
plot_time_series_regression(
    .data = data_prepared_tbl
    , .date_var = purchased_at
    , .formula = revenue ~ splines::ns(purchased_at, df = 4)
)
```


### LM Model Spec

Create a model specification for linear regression:

- Use `linear_reg()` function
- Use `set_engine("lm")`
- Store as `model_spec_lm`

```{r}
# YOUR CODE HERE
model_spec_lm <- linear_reg() %>%
    set_engine("lm")
```

### Recipe Spec - Spline

Create a recipe for the spline model. 

1. Start with `recipe_spec_base`
2. Add a step to remove the "purchased_at" feature. We don't need this for LM models. 
3. Add a step for the natural spline. Set `deg_free = 3`
4. Remove any features that begin with "lag_"
5. Store your updated recipe as `recipe_spec_1_spline`
6. Glimpse the output. Were the features adjusted correctly?

```{r}
# YOUR CODE HERE
recipe_spec_1_spline <- recipe_spec_base %>%
    step_rm(purchased_at) %>%
    step_ns(ends_with("index.num"), deg_free = 3) %>%
    step_rm(starts_with("lag_"))
```


### Workflow - Spline

Create a workflow for the linear regression and preprocessing recipe:

- Start with a `workflow()`
- Use `add_model()` to add the `model_spec_lm`
- Use `add_recipe()` to add the `recipe_spec_1_spline`
- Store as `workflow_fit_lm_1_spline`


```{r}
# YOUR CODE HERE
workflow_fit_lm_1_spline <- workflow() %>%
    add_model(model_spec_lm) %>%
    add_recipe(recipe_spec_1_spline) %>%
    fit(training(splits))
```

## Rolling Lag Model

### Recipe Spec - Lag

Create a recipe for the spline model. 

1. Start with `recipe_spec_base`
2. Add a step to remove the "purchased_at" feature. We don't need this for LM models. 
3. Remove missing values in any column that starts with "lag_"
4. Store your updated recipe as `recipe_spec_1_spline`
5. Glimpse the output. Were the features adjusted correctly?

```{r}
# YOUR CODE HERE
recipe_spec_2_spline <- recipe_spec_base %>%
    step_rm(purchased_at) %>%
    step_naomit(starts_with("lag_"))
```

### Workflow - Lag

Save the workflow as `workflow_fit_lm_2_lag`.

```{r}
# YOUR CODE HERE
workflow_fit_lm_2_lag <- workflow() %>%
    add_model(model_spec_lm) %>%
    add_recipe(recipe_spec_2_spline) %>%
    fit(training(splits))
```

# Modeltime

```{r}
# write_rds(workflow_fit_lm_1_spline, "challenge_02_data_checkpoints/workflow_fit_lm_1_spline.rds")
# write_rds(workflow_fit_lm_2_lag, "challenge_02_data_checkpoints/workflow_fit_lm_2_lag.rds")

# Checkpoint data
workflow_fit_lm_1_spline <- read_rds("challenge_02_data_checkpoints/workflow_fit_lm_1_spline.rds")
workflow_fit_lm_2_lag    <- read_rds("challenge_02_data_checkpoints/workflow_fit_lm_2_lag.rds")
```

### Make a Modeltime Table

Start by making a modeltime table:

- Use `modeltime_table()` to store your fitted workflows
- Save as `model_tbl`

```{r}
# YOUR CODE HERE
model_tbl <- modeltime_table(
    workflow_fit_lm_1_spline,
    workflow_fit_lm_2_lag
)
```

As a precautionary measure, please refit the models using `modeltime_refit()`. This prevents models that can go bad over time because of software changes. 

```{r}
# Refitting makes sure your models work over time. 
model_tbl <- model_tbl %>%
    modeltime_refit(training(splits))
```


### Calibrate the Table

Use testing data to calibrate the model:

- Start with `model_tbl`
- Use `modeltime_calibrate()` to calibrate the model using `testing(splits)` (out-of-sample data)
- Store the result as `calibration_tbl`

```{r}
# YOUR CODE HERE
calibration_tbl <- model_tbl %>%
    modeltime_calibrate(testing(splits))
```

### Calculate the Accuracy

Use `modeltime_accuracy()` to calculate the accuracy metrics.

```{r}
# YOUR CODE HERE
calibration_tbl %>%
    modeltime_accuracy()
```

### Visualize the Model Forecast

- Use `modeltime_forecast()`:
    - Set `new_data = testing(splits)`
    - Set `actual_data = data_prepared_tbl`
- Pipe the result into `plot_modeltime_forecast()`

```{r}
# YOUR CODE HERE
model_tbl %>%
    modeltime_forecast(
        new_data = testing(splits)
        , actual_data = data_prepared_tbl
    ) %>%
    plot_modeltime_forecast()
```


Forecasting thoughts:

- What can you say about the Rolling Lag Model?
- What might we be able to do to correct the model? (HINT: Try removing features in the lag model - what happens?)

# Forecast Future Data

```{r}
# write_rds(calibration_tbl, "challenge_02_data_checkpoints/calibration_tbl.rds")

# Checkpoint data
calibration_tbl <- read_rds("challenge_02_data_checkpoints/calibration_tbl.rds")
```


## Refit the Model

- Start with the `calibration_tbl`
- Use `modeltime_refit()` refit the model on the `data_prepared_tbl` dataset

```{r}
# YOUR CODE HERE
refit_tbl <- calibration_tbl %>%
    modeltime_refit(data = data_prepared_tbl)
```


## Forecast

1. Start with `refit_tbl`
2. Use `modeltime_forecast()` to forecast the `new_data = forecast_tbl`. Use `data_prepared_tbl` as the actual data. 
3. Plot the forecast using `plot_modeltime_forecast()`

```{r}
# YOUR CODE HERE
refit_tbl %>%
    modeltime_forecast(
        new_data = forecast_tbl
        , actual_data = data_prepared_tbl
    ) %>%
    plot_modeltime_forecast()
```

## Invert Transformation

Apply the inversion to the forecast plot:

- Invert the standardization
- Invert the log transformation

```{r}
# YOUR CODE HERE
refit_tbl %>%
    modeltime_forecast(
        new_data = forecast_tbl
        , actual_data = data_prepared_tbl
    ) %>%
    # invert transformation
    mutate(across(.value:.conf_hi, .fns = ~ standardize_inv_vec(
        x    = .,
        mean = std_mean,
        sd   = std_sd
    ))) %>%
    mutate(across(.value:.conf_hi, .fns = exp)) %>%
    plot_modeltime_forecast()
```




# Forecast Review

- What do we think of these forecasts?
- How can we further improve them? (HINT: what have competitors used that we haven't?)

