---
title: "Challenge 03: Revenue Forecasting with ARIMA, Prophet, & ETS/TBATS"
subtitle: "DS4B 203-R, Time Series Forecasting for Business"
author: "Business Science"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = T,
    warning = T,
    paged.print = FALSE, 
    # This should allow Rmarkdown to locate the data
    root.dir = rprojroot::find_rstudio_root_file()
)
```

# Challenge Objective

Your goal is to perform an 8-week revenue forecast using __ARIMA, Prophet, Exponential Smoothing, and TBATS models.__ You will create 11 models that experiment with different parameter settings and feature engineering. 


# Libraries

```{r, message=F, warning=F}
# Modeling
library(tidymodels)
library(modeltime)

# Core
library(tidyverse)
library(timetk)
library(lubridate)
```


# Picking Up from Challenge 02

We'll read in the artifacts that were produced in Challenge 02. 

## Challenge 2 Artifacts

First, import the `challenge_02_artifacts`, and let's recap where we are picking up from.

```{r}
challenge_02_artifacts  <- read_rds("challenge_03_data_checkpoints/challenge_02_artifacts.rds")
```

Interactively explore the artifacts to refresh your memory from Challenge 02. 

```{r, eval=FALSE}
challenge_02_artifacts %>% View("artifacts")
```


## Data Preparation

Load the artifacts. 

```{r}
# Processed Data
data_prepared_tbl <- challenge_02_artifacts$processed_data$data_prepared_tbl
forecast_tbl      <- challenge_02_artifacts$processed_data$forecast_tbl

# Train/Test Splits
splits            <- challenge_02_artifacts$train_test_splits$splits

# Linear Regression Model
workflow_fit_lm_1_spline <- challenge_02_artifacts$trained_models$workflow_fit_lm_1_spline

# Inversion Parameters
std_mean <- challenge_02_artifacts$transformation_params$standardize_params$std_mean
std_sd   <- challenge_02_artifacts$transformation_params$standardize_params$std_sd
```

# Train/Test Splits

The `splits` object contains your train/test split. We'll be modeling with it. 

- Extract the training data using `training()` function. 

```{r}
# CODE HERE
training(splits)
```

Visualize the train/test split.

- Use `tk_time_series_cv_plan()` to convert the `splits` object to a data frame that can be plotted
- Plot the train/test split using `plot_time_series_cv_plan()`

```{r}
# CODE HERE
tk_time_series_cv_plan(splits) %>%
    plot_time_series_cv_plan(purchased_at, revenue)
```


# Modeling 

Now that we have a feel for the training and testing data, let's dive into modeling. There are 3 Sections:

- __ARIMA:__ Models 1-5
- __Prophet:__ Models 6-9
- __Exponential Smoothing & TBATS:__  Models 10-11

## ARIMA

We'll start by modeling the revenue with ARIMA. 

### Model 1 - Basic Auto ARIMA

Let's start with a Basic Auto ARIMA:

- Start with `arima_reg()` to begin the specification
- Pipe into `set_engine()`. Specify "auto_arima" as the engine. 
- Pipe into `fit()`:
    - Use formula = `revenue ~ purchased_at`
    - Set `data = training(splits)`
- Store the output as `model_fit_1_arima_basic`
- Print the output and check out the parameters & AIC.

```{r, message = T}
# CODE HERE
model_fit_1_arima_basic <- arima_reg() %>%
    set_engine("auto_arima") %>%
    fit(revenue ~ purchased_at, data = training(splits))
```

Questions:
- What are the ARIMA orders? 0,1,1
    - How many lags were used? 1
    - Was differencing performed? Yes 1
    - Were any lagged error features used?
- Did modeltime auto-generate a frequency for ARIMA? Yes
- Is this a seasonal model? Yes

### Model 2 - Add Product Events

Next, repeat the model this time adding events:

- Modify the previous formula to include `event_november_sale` and `event_product_launch`
- Store the object as `model_fit_2_arima_xregs`
- Print the output to the screen. 

Questions:
- Has the AIC improved? 
- What coefficients do you see in your model?
- Is this a seasonal model?

```{r, message = T}
# CODE HERE
model_fit_2_arima_xregs <- arima_reg() %>%
    set_engine("auto_arima") %>%
    fit(revenue ~ purchased_at + event_november_sale + event_product_launch, 
        data = training(splits))
```

### Model 3 - Add Seasonality + Product Events

Use `plot_acf_diagnostics()` to inspect the ACF of your ARIMA model as follows:

- If a Difference was used in your ARIMA models, then apply differencing using `diff_vec()`

Question:
- Do you see any spikes in ACF Lags?
- Do you see any spikes in PACF Lags?

```{r, message = T}
# CODE HERE
plot_acf_diagnostics(
    .data = training(splits),
    .date_var = purchased_at,
    .value = diff_vec(revenue)
)
```


Make your 3rd Auto ARIMA model the same as your 2nd model (include events in your model formula). This time:

- Set `arima_reg(seasonal_period = 4)` to capture the ACF Lag 4 as a seasonality
- Store as `model_fit_3_arima_sarimax`

Question:
- Is this a seasonal model?


```{r}
# CODE HERE
model_fit_3_arima_sarimax <- arima_reg(seasonal_period = 4) %>%
    set_engine("auto_arima") %>%
    fit(revenue ~ purchased_at + event_november_sale + event_product_launch, 
        data = training(splits))

```

### Model 4 - Force Seasonality w/ Regular ARIMA

We can force seasonality using Regular (non-auto) ARIMA.

- Start with `arima_reg()`:
    - Set `seasonal_period = 4` to force a 4-period (monthly) seasonality
    - Set your non seasonal arima parameters to (2,1,2) 
    - Set your seasonal arima parameters to (1,0,1)
- Set your engine to `"arima"`
- Use the formula with events added. 
- Store your model as `model_fit_4_arima_sarimax`

Questions:
- Is this a seasonal model?
- Does adding the seasonality seem to improve the ARIMA model?

```{r}
# CODE HERE
model_fit_4_arima_sarimax <- arima_reg(
    seasonal_period = 4,
    non_seasonal_ar = 2,
    non_seasonal_differences = 1,
    non_seasonal_ma = 2,
    seasonal_ar = 1,
    seasonal_differences = 0,
    seasonal_ma = 1
) %>%
    set_engine("arima") %>%
    fit(revenue ~ purchased_at + event_november_sale + event_product_launch, 
        data = training(splits))
    
```

### Model 5 - Use Fourier Terms + Product Events instead of Seasonality

Last try to improve the Auto ARIMA model. Let's add a fourier series at period = 4 to try to capture the strong ACF in Lag 4.

- Start with `arima_reg()`
- Pipe into `set_engine()` using "auto_arima"
- Use a formula that includes events
- Add to your forumla a `fourier_vec()` at `period = 4`. 
- Store your model as `model_fit_5_arima_xreg_fourier`
- Print the model 

Questions:
- Does the added fourier term seem to improve the AIC of the model?

```{r}
# CODE HERE
model_fit_5_arima_xreg_fourier <- arima_reg() %>%
    set_engine("auto_arima") %>%
    fit(
        revenue ~ purchased_at + event_november_sale + event_product_launch +
            fourier_vec(revenue, 4),
        data = training(splits)
    )
```

### Investigate - Modeltime Workflow

Next, we need to investigate our ARIMA models. 

#### Model Table

Use `modeltime_table()` to consolidate Models 1 - 5 (ARIMA models). Save the modeltime table as `model_tbl_arima`.

```{r}
# CODE HERE
model_tbl_arima <- modeltime_table(
    model_fit_1_arima_basic,
    model_fit_2_arima_xregs,
    model_fit_3_arima_sarimax,
    model_fit_4_arima_sarimax,
    model_fit_5_arima_xreg_fourier
)
```


#### Calibration Table

Use `modeltime_calibrate()` to calibrate your models on the testing split. 

```{r}
# CODE HERE
calibration_tbl <- model_tbl_arima %>%
    modeltime_calibrate(testing(splits))
```

#### Test Accuracy

Use `modeltime_accuracy()` to calculate the test accuracy. 

```{r}
# CODE HERE
calibration_tbl %>%
    modeltime_accuracy()
```

#### Test Forecast

Make a test forecast using `modeltime_forecast()` and `plot_modeltime_forecast()`. Use the testing split and the `data_prepared_tbl`.

```{r}
# CODE HERE
calibration_tbl %>%
    modeltime_forecast(
        new_data = testing(splits),
        actual_data = data_prepared_tbl
    ) %>%
    plot_modeltime_forecast()
```

### ARIMA Forecast Review

- Which ARIMA forecasts performed the best on the test data set?
    - Which ARIMA forecasts had the lowest MAE / RMSE?
    - Which ARIMA forecasts had the highest variance explained
- How did the forecasts do predicting the global trend?
- How did the forecasts do predicting the local trend?

```{r}
# Checkpoint data
# model_tbl_arima %>% write_rds("challenge_03_data_checkpoints/model_tbl_arima.rds")
model_tbl_arima <- read_rds("challenge_03_data_checkpoints/model_tbl_arima.rds")
model_tbl_arima
```



## Prophet 

Next, let's experiment with the prophet algorithm.

### Model 6 - Basic Prophet 

First, train a basic `prophet` model:

- Start with `prophet_reg()`. Use the default parameters. 
- Set the engine to "prophet"
- Fit the model to the training data using `revenue ~ purchased_at`
- Store the model as `model_fit_6_prophet_basic`


```{r}
# CODE HERE
```

### Model 7 - Turn on yearly seasonality

Next, let's try toggling yearly seasonality on. Make a new model called `model_fit_7_prophet_yearly` with `seasonality_yearly = TRUE`.

```{r}
# CODE HERE
```

### Model 8 - Product Events

Let's try one without yearly seasonality but now adding events. Make `model_fit_8_prophet_events` using the default settings for `prophet_reg()` and updating the fitting formula to include `event_november_sale` and `event_product_launch`. 

```{r}
# CODE HERE
```

### Model 9 - Events + Fourier Series

Let's try one more model that includes the events and a fourier series. Add a `fourier_vec()` with `period = 4`. Save this model as `model_fit_9_prophet_events_fourier`.

```{r}
# CODE HERE
```

### Investigate - Modeltime Workflow

Now let's check out the results. 

#### Model Table

Create a modeltime table with each of the prophet models 6-9 in the table. Store as `model_tbl_prophet`.

```{r}
# CODE HERE
```


#### Calibration Table

Next, calibrate the models using your testing set. 

```{r}
# CODE HERE
```

#### Test Accuracy

Calculate the accuracy with `modeltime_accuracy()`.

```{r}
# CODE HERE
```

#### Test Forecast

Finally, forecast the calibrated models on the testing data using `modeltime_forecast()` and `plot_modeltime_forecast()`.

```{r}
# CODE HERE
```

### Prophet Forecast Review

- Which Prophet forecasts performed the best on the test data set?
    - Which Prophet forecasts had the lowest MAE / RMSE?
    - Which Prophet forecasts had the highest variance explained
- How did the forecasts do predicting the global trend?
- How did the forecasts do predicting the local trend?

```{r}
# Checkpoint data
# model_tbl_prophet %>% write_rds("challenge_03_data_checkpoints/model_tbl_prophet.rds")
model_tbl_prophet <- read_rds("challenge_03_data_checkpoints/model_tbl_prophet.rds")
model_tbl_prophet
```

## Exponential Smoothing

Now let's try models that incorporate exponential smoothing. 

### Model 10 - ETS

The first model we'll experiment with is the automated "ETS" model:

- Start with `exp_smoothing()`
- Set the engine to "ets"
- Fit the model to the training data using a formula `revenue ~ purchased_at`
- Save the model as `model_fit_10_ets`

Question:

- Review the ETS parameters:
    - Is this an exponentially smoothed error model?
    - Is this a trend model
    - Is this a seasonal model?
- How does the AIC compare to the AIC of the ARIMA model?


```{r}
# CODE HERE
```



### Model 11 - TBATS

Next, let's make a TBATS model. The seasonality we'll use is `seasonal_period_1 = 4` and `seasonal_period_2 = 13`.

```{r}
# CODE HERE
```


### Investigate - Modeltime  Workflow

Now let's check out the results. 

#### Model Table

Create a modeltime table with each of the exponential smoothing models 10-11 in the table. Store as `model_tbl_exp_smooth`.

```{r}
# CODE HERE
```

#### Calibration Table

Next, calibrate the models using your testing set. 

```{r}
# CODE HERE
```

#### Test Accuracy

Calculate the accuracy with `modeltime_accuracy()`.

```{r}
# CODE HERE
```

#### Test Forecast

Finally, forecast the calibrated models on the testing data using `modeltime_forecast()` and `plot_modeltime_forecast()`.

```{r}
# CODE HERE
```
### Exponential Smoothing Forecast Review

- Which Exponential Smoothing forecasts performed the best on the test data set?
    - Which Exponential Smoothing forecasts had the lowest MAE / RMSE?
    - Which Exponential Smoothing forecasts had the highest variance explained
- How did the forecasts do predicting the global trend?
- How did the forecasts do predicting the local trend?

```{r}
# Checkpoint data
# model_tbl_exp_smooth %>% write_rds("challenge_03_data_checkpoints/model_tbl_exp_smooth.rds")
model_tbl_exp_smooth <- read_rds("challenge_03_data_checkpoints/model_tbl_exp_smooth.rds")
model_tbl_exp_smooth
```



# Forecast Future Data

Forecast the future.

```{r}
# Checkpoint data
model_tbl_arima      <- read_rds("challenge_03_data_checkpoints/model_tbl_arima.rds")
model_tbl_prophet    <- read_rds("challenge_03_data_checkpoints/model_tbl_prophet.rds")
model_tbl_exp_smooth <- read_rds("challenge_03_data_checkpoints/model_tbl_exp_smooth.rds")
```


### Model Table

Create a Modeltime Table from the 3 previous Modeltime Tables using a new function, `combine_modeltime_tables`.

Use `?combine_modeltime_tables` to learn about how the function works.

```{r, eval=FALSE}
?combine_modeltime_tables
```


Use the function to combine the 3 previous modeltime tables into a single modeltime table. Combine these Modeltime Tables: 
    - `model_tbl_arima`
    - `model_tbl_prophet`
    - `model_tbl_exp_smooth`
Store the combined modeltime tables as `model_tbl`

```{r}
# CODE HERE
```

As a precautionary measure, please refit the models using `modeltime_refit()`. This prevents models that can go bad over time because of software changes. 

```{r}
# Refitting makes sure your models work over time. 
model_tbl <- model_tbl %>%
    modeltime_refit(training(splits))
```

### Calibrate the Table

Use testing data to calibrate the model:

- Start with `model_tbl`
- Use `modeltime_calibrate()` to calibrate the model using `testing(splits)` (out-of-sample data)
- Store the result as `calibration_tbl`

```{r}
# CODE HERE
```

### Calculate the Accuracy

Use `modeltime_accuracy()` to calculate the accuracy metrics.

```{r}
# CODE HERE
```

### Visualize the Model Forecast

- Use `modeltime_forecast()`:
    - Set `new_data = testing(splits)`
    - Set `actual_data = data_prepared_tbl`
- Pipe the result into `plot_modeltime_forecast()`

```{r}
# CODE HERE
```



### Refit

- Start with the `calibration_tbl`
- Use `modeltime_refit()` refit the model on the `data_prepared_tbl` dataset

```{r}
# CODE HERE
```

### Forecast Future 

1. Start with `refit_tbl`
2. Use `modeltime_forecast()` to forecast the `new_data = forecast_tbl`. Use `data_prepared_tbl` as the actual data. 
3. Plot the forecast using `plot_modeltime_forecast()`

```{r}
# CODE HERE
```



## Invert Transformation

Apply the inversion to the forecast plot:

- Invert the standardization
- Invert the log transformation


```{r}
# CODE HERE
```

# Forecast Review

- What do we think of these forecasts?
- Which models had the best performance? Are there any similarities?
- Were any forecasts inconsistent with your expectations? If so, why might this be?

